On a Alfvenicity of solar wind discontinuities

### 2024-09-30 article.qmd
Update multi-fluid model for force-free current sheet equilibria

|      | Harris Current Sheet                              | Yoon-Lui Current Sheet                               | Kan Current Sheet                        |
|-------------------|-------------------------------------------------------|--------------------------------------------------------|-----------------------------------------------------|
| **Magnetic Field** | $B_x(z) = B_0 \tanh(z/L)$ | $B_x(z) = B_0 \tanh \left(\frac{h}{L} \tan^{-1}(z/h)\right)$|$B_x^*(x^*, z^*)$ |
| **Number Density** | $n(z) = n_0 \text{sech}^2(z/L)$ | $n(z) = n_0 \text{sech}^2 \left(\frac{h}{L} \tan^{-1}(z/h)\right)$ | $n^*(x^*, z^*)$ |
| **Current Density** | $j_y(z) = B_0 \text{sech}^2(z/L) / (\mu_0 L)$ |$j_y(z) = \frac{B_0}{\mu_0 L} \frac{h^2}{z^2 + h^2} \text{sech}^2 \left(\frac{h}{L} \tan^{-1}(z/h)\right)$|                     |
| **Drift Velocity** |  |$V_y(z) = V_0 \frac{h^2}{(z^2 + h^2)}$ |                                       |
| **Special Feature** | Vlasov equilibrium, Maxwellian velocity distribution | Velocity shear important for thin current sheets       | Function form involves complex variables            |

@@

::: {#fig-examples}

![](figures/fig_examples.svg)

Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.

:::

@@

We used a steady-state multi-fluid model for assumed 1-D force-free current sheets. This approach, inspired by @steinhauerMultifluidModelOnedimensional2008, has the advantage of being analytically tractable while providing the necessary complexity to include the kinetic effects of multiple ion species. Kinetic approach, which is in principle more accurate, is hard to solve analytically and difficult to interpret the measurements. The multifluid model is a good compromise between the simplicity of the MHD model and the complexity of the kinetic model. One disadvantage of the multifluid model, however, is that there are more unknowns than equations. This means that the model is underdetermined and requires additional assumptions to be solved. This closure problem is usually solved by choosing equations of state. In this study, we first attempted to solve the system with a scalar pressure model, but then the system was overdetermined and no solution was found. We avoid this problem by assuming a specific profile for one of the variables, which is then used to solve the rest of the system.

@@

@@

$$
\begin{aligned}
Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
\\
Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
\\
Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
\end{aligned}
$${#eq-momentum}

@@

$$
d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x+\Gamma_e B_x / B_z
$${#eq-Jx}

$$
d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y-\Gamma_e B_y / B_z
$${#eq-Jy}

$$
d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z-\Gamma_e
$${#eq-Jz}

@@

$$
n_α (u_{αx} B_y - n_α u_{αy} B_x) = 0
$${#eq-alpha}

@@

Conveniently, the center of the current sheet is chosen as the origin $z = 0$, which corresponds to the lower boundary of @eq-theta1. As a result, the boundary condition for the rotation angle is given by  $\theta_\alpha(0) = \pi/2$.

@@

::: {#fig-profiles layout-ncol="2" }

![](figures/profiles_sym.svg){#fig-profilesEx1}

@@

(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.

:::

@@

::: {#fig-profilesEx2 layout-ncol="2" }

![](figures/profiles_n1Inf=0.6.svg)

@@

Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.

:::

@@

::: {#fig-profilesC layout-nrow="1" }

![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}

![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}

:::

>>>>

@@

::: {#fig-examples}
![](figures/fig_examples.svg)

Three examples of solar wind discontinuities observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.
:::

@@

We used a multi-fluid model for assumed 1-D quasineutral force-free current sheet equilibria. This approach, inspired by @steinhauerMultifluidModelOnedimensional2008, has the advantage of being analytically tractable while providing the necessary complexity to include the kinetic effects of multiple ion species. Kinetic approach, which is in principle more accurate, is hard to solve analytically and difficult to interpret the measurements. The multifluid model is a good compromise between the simplicity of the MHD model and the complexity of the kinetic model. One disadvantage of the multifluid model, however, is that there are more unknowns than equations. This means that the model is underdetermined and requires additional assumptions to be solved. This closure problem is usually solved by choosing equations of state. In this study, we first attempted to solve the system with a scalar pressure model, but then the system was overdetermined and no solution was found. We avoid this problem by assuming a specific profile for one of the variables, which is then used to solve the rest of the system.

In this model, we assume that the asymptotic boundary conditions are known from measurements, including the normal component of the magnetic field and the asymptotic properties of the incident fluid streams, which are clearly distinguishable. It should be noted, however, that while the velocity space boundary between different groups may be clear in the uniform field region outside the current sheet, within the sheet itself, the boundary can become more complex. For instance, individual ions may jump from one group to another, and some may become trapped within the current sheet. These complexities, particularly ion behavior within the sheet, are beyond the scope of this paper and are not addressed in the current analysis.

@@

The governing equations for the multi-fluid plasma model can be expressed in conservation law form as follows:

$$
(\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) n_α = - n_α ∇ · \mathbf{u}_α
$$

$$
m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right) + \mathbf{R}_α
$$

where $α$ indicates the particle species, $n_α$ the number density, $m_α$ the mass, $q_α$ the charge, $\mathbf{u}_α$ the bulk velocity, $\mathbf{P}_α$ the pressure tensor, $\mathbf{E}$ the electric field, $\mathbf{B}$ the magnetic induction (magnetic field), and $R_α$ the friction forces.

In this study, we focus on the equilibrium solutions of the set of equations, which implies that the partial derivatives with respect to time are set to zero ($\frac{∂}{∂ t}=0$). 
Our primary interest lies in the current sheet, and we simplify the problem by assuming that the magnetic field depends only on a single Cartesian coordinate, designated as $z$.
Additionally, we assume that the background electron is massless, meaning the electrons follow the magnetic field lines $\mathbf{u}_e \times \mathbf{B} = 0$. 
For further simplification, we also assume the ion populations are gyrotropic, eliminating any non-diagonal terms in the pressure tensor.

@@

$$
\begin{aligned}
Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
\\
Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
\\
Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
\end{aligned}
$$ {#eq-momentum}

@@

$$
d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
$$ {#eq-Jx}

$$
d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
$$ {#eq-Jy}

$$
d B_z / d z = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
$$ {#eq-Jz}

@@

$$
n_α (u_{αx} B_y - n_α u_{αy} B_x) = 0
$$ {#eq-alpha}

@@

Conveniently, the center of the current sheet is chosen as the origin $z = 0$, which corresponds to the lower boundary of @eq-theta1. As a result, the boundary condition for the rotation angle is given by  $\theta_α(0) = \pi/2$.

@@

::: {#fig-profiles layout-ncol="2" }
![](figures/profiles_sym.svg){#fig-profilesEx1}

@@

(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.
:::

@@

::: {#fig-profilesEx2 layout-ncol="2" }
![](figures/profiles_n1Inf=0.6.svg)

@@

Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
:::

@@

::: {#fig-profilesC layout-nrow="1" }
![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}

![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}
:::
